### 2015-01-07 prelim/oral_prelim.tex
Oral prelim updates

\documentclass{beamer}
\usepackage{beamerthemesplit}
% \usepackage{pstricks}
\usepackage{graphicx}
\usepackage{mdwlist}
\usepackage{lineno, hyperref}

@@

\title[Spatial methods for extreme value analysis] % (optional, use only with long paper titles)
{
  Spatial methods for extreme value analysis
}
\author[Samuel A. Morris]{Samuel A. Morris}
\institute[NCSU]{}
\date{January 9, 2015}

@@

\begin{frame}{Motivation}
  \centering
  \begin{figure}
    \includegraphics[width=\linewidth, trim=0 1in 0 0in ]{./plots/pot/ozone-10jul-us.pdf}
    \caption{Max 8-hour ozone measurements on July 10, 2005}
   \end{figure}
\end{frame}

@@

  \column{.55\linewidth}
  \begin{figure}
    \includegraphics[width=1\linewidth]{./plots/define_extreme.pdf}
    \caption{Monthly maximums recorded over a three years.}
    \end{figure}
\end{columns}
\end{frame}

\begin{frame}{Standard analysis - Block maxima}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Asymptotic result:
    \begin{itemize}
      \item Let $X_1, \ldots, X_n$ be i.i.d.
      \item Consider $M_n = \max(X_1, \ldots, X_n)$
      \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
      \begin{align*}
        a_n^{-1} (M_n - b_n) \converged G(z)
      \end{align*}
      then $G(z)$ follows a generalized extreme value distribution
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard analysis - Block maxima}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Generalized extreme value distribution has three parameters:
    \begin{itemize}
      \item $\mu \in \calR$ is a location parameter
      \item $\sigma > 0$ is a scale parameter
      \item $\xi \in \calR$ is a shape parameter
      \begin{itemize}
        \item Unbounded above if $\xi \ge 0$
        \item Bouded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
      \end{itemize}
    \end{itemize}
    \item Challenge:
    \begin{itemize}
      \item Lose information by only considering maximum in a block
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard analysis - Block maxima}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Generalized extreme value distribution
    \begin{align*}
      G(y) = \Pr(Y < y) = \left\{  \begin{array}{ll}
        \exp\left\{ -\left[ 1 + \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{ -1 / \xi} \right\} & \quad \xi \neq 0 \\[0.5em]
        \exp \left\{ -\exp \left( - \frac{ y - \mu }{ \sigma} \right) \right\} & \quad \xi = 0
      \end{array}\right.
    \end{align*}
    \item Standardized distribution is unit Fr\'{e}chet or GEV(1, 1, 1)
    \begin{align*}
      \Pr(Z < z) = exp(-z^{-1})
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Standard analysis - Peaks over threshold}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Generalized Pareto distribution has two parameters:
    \begin{itemize}
      \item $\sigma > 0$ is a scale parameter
      \item $\xi \in \calR$ is a shape parameter
      \begin{itemize}
        \item Unbounded above if $\xi \ge 0$
        \item Bouded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
      \end{itemize}
    \end{itemize}
    \item Challenges:
    \begin{itemize}
      \item Sensitive to threshold selection
      \item Temporal dependence between observations (e.g. flood levels don't dissipate overnight)
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard analysis - Peaks over threshold}
  \begin{itemize}  \setlength{\itemsep}{0.5em}
    \item Select a threshold, $T$, and use the Generalized Pareto distribution to model the exceedances
    \begin{align*}
      H(y) = P(Y < y) = \left\{ \begin{array}{ll}
        1 - \left[1 - \xi \left( \frac{ y - T }{ \sigma } \right) \right]^{-1 / \xi} & \quad \xi \neq 0 \\[0.5em]
        1 - \exp \left\{ \frac{ y - T }{ \sigma} \right\} & \quad \xi = 0
      \end{array}\right.
    \end{align*}
    \item Related to GEV distribution through
    \begin{align*}
      H(y) = 1 + \log[G(y)], \quad y \ge T
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Max-stable processes}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution (Cooley et al., 2012):
    \begin{itemize}
      \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$.
      \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
      \item If there exists normalizing sequences $a_T(\bs)$ and $b_T(\bs)$ such
      that for all sites, $\bs_i, i = 1, \ldots, d$,
      \begin{align*}
        a_T^{-1}(\bs) \left\{ M_T(\bs) - b_T(\bs) \right\} \converged Y(\bs)
      \end{align*}
      which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate representations}
  \begin{itemize}
    \item Marginally at each site, observations follow a generalized extreme value distribution
    \item Finite collection of sites:
    \begin{itemize}
      \item The multivariate representation for the GEV is
      \begin{align*}
        \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
                V(\bs)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
      \end{align*}
      where
      \begin{itemize}
        \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
        \item $H$ is a probability measure on $\Delta_d$
        \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Extremal dependence: $\chi$ statistic}
  \begin{itemize} \setlength{\itemsep}{0.5em}
   \item The $\chi$ statistic is a measure of extremal dependence in the tails
   \item Specifically, we focus on $\chi(h)$ for the upper tail given by
    \begin{align*}
      \chi(h) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bt) > c)
    \end{align*}
    where $h = ||\bs - \bt||$
    \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $h$
  \end{itemize}
\end{frame}

\begin{frame}{Existing challenges}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Multivariate max-stable and GPD models have nice features, but they are
    \begin{itemize}
      \item computationally challenging (Falk et al., 2011)
      \begin{itemize}
        \item Asymmetric logistic has $2^{n-1}(n + 2) - (2n + 1)$ free parameters
      \end{itemize}
      \item joint distribution only available in low dimension
    \end{itemize}
    \item Some recent approaches:
    \begin{itemize}
      \item Bayesian hierarchical model (Reich and Shaby, 2012)
      \item Pairwise likelihood approach (Huser and Davison, 2014)
    \end{itemize}
    \item Many opportunities to explore new methods
  \end{itemize}
\end{frame}

\begin{frame}{Three principal contributions}
  \begin{enumerate}[1] \setlength{\itemsep}{0.5em}
    \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
    \item Predicting exceedances using a spatially dependent generalized extreme-value link function
    \item A Bayesian hierarchical model to allow for non-stationary covariance in extreme value models.
  \end{enumerate}
\end{frame}

\begin{frame}{Spatiotemporal modeling for extreme values}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Model to analysis spatiotemporal extreme values
    \item Model objectives:
    \begin{itemize}
      \item Has marginal distribution with a flexible tail
      \begin{itemize}
        \item allow for asymmetric distributions
        \item allow for heavy tails
      \end{itemize}
      \item Has asymptotic spatial dependence
      \item Has computation on the order of Gaussian models for large space-time datasets
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Gaussian spatial model}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
    \item Model properties:
    \begin{itemize}
      \item Nice computing properties (closed-form likelihood)
      \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
      \item Tail is not flexible:
      \begin{itemize}
        \item light tailed
        \item symmetric
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
    \begin{align*}
      Y(\bs) \sim \text{skew-t}(\bX(\bs), \bOmega, \alpha, \text{df}=2a)
    \end{align*}
    where
    \begin{itemize}
      \item $\bX(\bs) \bbeta$ is the location
      \item $\bOmega = \frac{1}{ab} \bar{\bOmega}$ is a correlation matrix
      \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
      \item $\Sigma$ is a postive definite correlation matrix
      \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{$\chi(h)$ plot}
  \vspace{-2em}
  \centering
  \begin{figure}
  \includegraphics[width=.65\linewidth]{./plots/pot/chi-h-t_and_gaus.pdf}\\[-0.2in]
  \caption{$\chi$ plot for symmetric $t$, skew-$t$, and Gaussian}
  \end{figure}
\end{frame}

\begin{frame}{Censoring data to focus on tail behavior}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item We censor the observed data at a high threshold $T$.
    \item Censored data:
    \begin{align*}
      \tilde{Y}_t(\bs) = \left\{ \begin{array}{ll}
          Y_t(\bs) \quad & \delta(\bs) = 1\\
          T \quad & \delta(\bs) = 0
      \end{array}\right.
    \end{align*}
    where $\delta(\bs) = I[Y(\bs) > T]$
    \item Allows tails of the distribution to speak for themselves.
  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution properties}
  \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Model properties
    \begin{itemize}
    	\item Has flexible tail:
      \begin{itemize}
        \item Skewness controlled by $\lambda$
        \item Weight of tails controlled by degrees of freedom $2a$
      \end{itemize}
    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011)
    	\item Computation that is on the order of Gaussian computation
    \end{itemize}
    \item Challenge: $\chi(h) > 0$ as $h \rightarrow \infty$ (Padoan, 2011)
    \begin{itemize}
      \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
      \item We deal with this through a daily random partition (similar to Kim et al., 2005)
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Example partition}
    \centering
    \begin{figure}
    \includegraphics[width=0.54\linewidth]{./plots/pot/example-partition-1.pdf}
    \includegraphics[width=0.54\linewidth]{./plots/pot/example-partition-2.pdf}
    \caption{Two sample partitions (number is at partition center)}
    \end{figure}
\end{frame}

\begin{frame}{$\chi(h)$ plot}
  \vspace{-2em}
  \centering
  \begin{figure}
  \includegraphics[width=.7\linewidth]{./plots/pot/chi-h.pdf}\\[-0.4in]
  \caption{$\chi$ plot for different data settings}
  \end{figure}
\end{frame}

\begin{frame}{Random partition skew-$t$ model}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item This new model is called a random partition skew-$t$ model, and it has the properties we desire
    \begin{itemize}
      \item Marginal distribution with flexible tails
      \begin{itemize}
        \item $\lambda$ term allows for asymmetry
        \item d.f. controls heavy vs light tails
      \end{itemize}
      \item Asymptotic spatial dependence for that decays with distance between sites via partition
      \item Computation is on the order of Gaussian models for large space-time datasets
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Temporal dependence}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Temporal dependence should be accounted for when using daily data
    \item For multiple days of observations the model becomes
      \begin{align*}
        Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) | z_t(\bs) | + \sigma_t(\bs) v_t(\bs)
      \end{align*}
      where $t = 1, \ldots, T$ denotes the day of each observation.
    \item Different ways to incorporate the temporal dependence
    \begin{itemize}
      \item Time series on $\bw_t$, $z_t(\bs)$, and $\sigma_t(\bs)$
      \item Three dimensional covariance model for $v_t(\bs)$ (e.g. Huser and Davison, 2014)
    \end{itemize}
    \item Our model incorporates the time series because the $z_t(\bs)$ and $\sigma_t(\bs)$ terms dictate the tail behavior.
  \end{itemize}
\end{frame}

@@

\begin{frame}{MCMC details}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Three main steps:
    \begin{enumerate}[1.]
      \item Impute censored data below $T$
      \item Update parameters with standard random walk Metropolis Hastings or Gibbs sampling
      \item Make spatial predictions
    \end{enumerate}
    \item Priors are selected to be conjugate when possible
  \end{itemize}
\end{frame}

@@

@@

\begin{frame}{Simulation study results}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Key findings:
    \begin{itemize}
      \item Improvement over Gaussian methods when partitioning
      \item Underestimating the number of knots has a detrimental impact
      \item In all cases, non-thresholded models perform better than thresholded models
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Data analysis}
\begin{columns}[c]
\column{.45 \linewidth}
	\begin{itemize} \setlength{\itemsep}{0.5em}
	\item Ozone measurements
	\begin{itemize}
		\item max 8-hour ozone measurements
		\item data from 1089 sites
		\item July 2005
	\end{itemize}
  \item We take a stratified sample of $n = 800$ sites:
  \begin{itemize}
    \item 271 from northeast
    \item 96 from northwest
    \item 269 from southeast
    \item 164 from southwest
  \end{itemize}
	\end{itemize}

@@

\begin{frame}{Data analysis}
  \centering
  \begin{figure}
    \includegraphics[width=\linewidth, trim=0 1in 0 0in ]{./plots/pot/ozone-10jul-us.pdf}
    \caption{Max 8-hour ozone measurements on July 10, 2005}
   \end{figure}
\end{frame}

\begin{frame}{Exploratory data analysis}
	\centering
  \begin{figure}
    \includegraphics[width=1\linewidth]{./plots/pot/chi-plot-ozone-res.pdf}
    \caption{$\widehat{\chi}$-plot for sample quantiles of ozone observations}
  \end{figure}
\end{frame}

>>>>

\documentclass{beamer}
% \usepackage{beamerthemesplit}
\usetheme{Madrid}
% \usepackage{pstricks}
\usepackage{graphicx}
\usepackage{mdwlist}
\usepackage{lineno, hyperref}
\usepackage{stackrel}

@@

\title[Spatial methods for EVA] % (optional, use only with long paper titles)
{
  Spatial methods for extreme value analysis
}
\author[S. Morris]{Samuel A. Morris}
\institute[]{North Carolina State University}
\date[]{January 9, 2015}

@@

\begin{frame}{Motivation}
  \centering
  \begin{figure}
    \includegraphics[width=\linewidth, trim=0 1in 0 1in ]{./plots/pot/ozone-10jul-us.pdf}
    \caption{Max 8-hour ozone measurements on July 10, 2005}
   \end{figure}
\end{frame}

\begin{frame}{Motivation}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Ozone compliance for Clean Air Act (EPA)
    \begin{itemize}
      \item Annual fourth-highest daily maximum 8-hour concentration, averaged over 3 years, not to exceed 75 ppb
      \item Annual fourth-highest is the 99th percentile for the year
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \column{.55\linewidth}
  \begin{figure}
    \includegraphics[width=1\linewidth, trim=0 0.5in 0 1in]{./plots/define_extreme.pdf}
    \caption{Monthly maximums recorded over a three years}
    \end{figure}
\end{columns}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Block maxima}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Asymptotic result (Falk et al., 2011)
    \begin{itemize}
      \item Let $X_1, \ldots, X_n$ be i.i.d.
      \item Consider $M_n = \max(X_1, \ldots, X_n)$
      \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
      \begin{align*}
        a_n^{-1} (M_n - b_n) \converged G(z)
      \end{align*}
      then $G(z)$ follows a generalized extreme value distribution
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Block maxima}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Generalized extreme value distribution has three parameters:
    \begin{itemize}
      \item $\mu \in \calR$ is a location parameter
      \item $\sigma > 0$ is a scale parameter
      \item $\xi \in \calR$ is a shape parameter
      \begin{itemize}
        \item Unbounded above if $\xi \ge 0$
        \item Bouded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
      \end{itemize}
    \end{itemize}
    \item Challenge:
    \begin{itemize}
      \item Lose information by only considering maximum in a block
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Block maxima}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Generalized extreme value distribution
    \begin{align*}
      G(y) = \Pr(Y < y) = \left\{  \begin{array}{ll}
        \exp\left\{ -\left[ 1 + \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{ -1 / \xi} \right\} & \quad \xi \neq 0 \\[0.5em]
        \exp \left\{ -\exp \left( - \frac{ y - \mu }{ \sigma} \right) \right\} & \quad \xi = 0
      \end{array}\right.
    \end{align*}
    \item Standardized distribution is unit Fr\'{e}chet or GEV(1, 1, 1)
    \begin{align*}
      \Pr(Z < z) = exp(-z^{-1})
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Asymptotic result (Falk et al., 2011)
    \begin{itemize}
      \item Let $X_1, \ldots, X_n \sim F$
      \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ with
      \begin{align*}
        \lim_{n \rightarrow \infty} 1 - F(b_n) = 0 \qquad \text{and} \qquad \lim_{ n \rightarrow \infty} \frac{ 1 - F(b_{n+1})}{1 - F(b_n)} = 1
      \end{align*}
      where $b_n$ is a sequence of thresholds
      \item If for any $x \ge 0$,
      \begin{align*}
        1 - \frac{ 1 - F(a_n x + b_n) }{ 1 - F(b_n) } \converged H(x),
      \end{align*}
      then $H(x)$ follows a generalized Pareto distribution
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Generalized Pareto distribution has two parameters:
    \begin{itemize}
      \item $\sigma > 0$ is a scale parameter
      \item $\xi \in \calR$ is a shape parameter
      \begin{itemize}
        \item Unbounded above if $\xi \ge 0$
        \item Bounded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
      \end{itemize}
    \end{itemize}
    \item Challenges:
    \begin{itemize}
      \item Sensitive to threshold selection
      \item Temporal dependence between observations (e.g. flood levels don't dissipate overnight)
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
  \begin{itemize}  \setlength{\itemsep}{0.5em}
    \item Select a threshold, $T$, and use the Generalized Pareto distribution to model the exceedances
    \begin{align*}
      H(y) = P(Y < y) = \left\{ \begin{array}{ll}
        1 - \left[1 - \xi \left( \frac{ y - T }{ \sigma } \right) \right]^{-1 / \xi} & \quad \xi \neq 0 \\[0.5em]
        1 - \exp \left\{ \frac{ y - T }{ \sigma} \right\} & \quad \xi = 0
      \end{array}\right.
    \end{align*}
    \item Related to GEV distribution through
    \begin{align*}
      H(y) = 1 + \log[G(y)], \quad y \ge T
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Max-stable processes}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution (Cooley et al., 2012)
    \begin{itemize}
      \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$
      \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
      \item If there exists normalizing sequences $a_T(\bs)$ and $b_T(\bs)$ such
      that for all sites, $\bs_i, i = 1, \ldots, d$,
      \begin{align*}
        a_T^{-1}(\bs) \left\{ M_T(\bs) - b_T(\bs) \right\} \converged Y(\bs)
      \end{align*}
      which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate representations}
  \begin{itemize}
    \item Marginally at each site, observations follow a generalized extreme value distribution
    \item For a finite collection of sites
    \begin{itemize}
      \item The multivariate representation for the GEV is
      \begin{align*}
        \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
                V(\bs)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
      \end{align*}
      where
      \begin{itemize}
        \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
        \item $H$ is a probability measure on $\Delta_d$
        \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Extremal dependence: $\chi$ statistic}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Correlation is the most common measure of dependence, but it is irrelevant for extreme value analysis
    \item Extreme value analysis focuses on the $\chi$ statistic, a measure of extremal dependence
    \item Specifically, we focus on $\chi(h)$ for the upper tail given by
    \begin{align*}
      \chi(h) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bt) > c)
    \end{align*}
    where $h = ||\bs - \bt||$
    \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $h$
  \end{itemize}
\end{frame}

\begin{frame}{Existing challenges}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Multivariate max-stable and GPD models have nice features, but they are
    \begin{itemize}
      \item Computationally challenging (Falk et al., 2011)
      \begin{itemize}
        \item Asymmetric logistic has $2^{n-1}(n + 2) - (2n + 1)$ free parameters
      \end{itemize}
      \item Joint distribution only available in low dimension
    \end{itemize}
    \item Some recent approaches
    \begin{itemize}
      \item Bayesian hierarchical model (Reich and Shaby, 2012)
      \item Pairwise likelihood approach (Huser and Davison, 2014)
    \end{itemize}
    \item Many opportunities to explore new methods
  \end{itemize}
\end{frame}

\begin{frame}{Three principal contributions}
  \begin{enumerate}[1] \setlength{\itemsep}{0.5em}
    \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
    \item Predicting rare binary events with a spatially dependent generalized extreme-value link function
    \item A Bayesian hierarchical model to allow for non-stationary covariance in extreme value models
  \end{enumerate}
\end{frame}

\begin{frame}{Spatiotemporal modeling for extreme values}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Model to analysis spatiotemporal extreme values
    \item Model objectives
    \begin{itemize}
      \item Has marginal distribution with a flexible tail
      \begin{itemize}
        \item Allow for asymmetric distributions
        \item Allow for heavy tails
      \end{itemize}
      \item Has asymptotic spatial dependence
      \item Has computation on the order of Gaussian models for large space-time datasets
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Gaussian spatial model}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
    \item Model properties
    \begin{itemize}
      \item Nice computing properties (closed-form likelihood)
      \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
      \item Tail is not flexible
      \begin{itemize}
        \item Light-tailed
        \item Symmetric
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
    \begin{align*}
      Y(\bs) \sim \text{skew-t}(\bX(\bs), \bOmega, \alpha, \text{df}=2a)
    \end{align*}
    where
    \begin{itemize}
      \item $\bX(\bs) \bbeta$ is the location
      \item $\bOmega = \frac{1}{ab} \bar{\bOmega}$ is a correlation matrix
      \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
      \item $\bSigma$ is a postive definite correlation matrix
      \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{$\chi(h)$ plot}
  \vspace{-2em}
  \centering
  \begin{figure}
  \includegraphics[width=.65\linewidth]{./plots/pot/chi-h-t_and_gaus.pdf}\\[-0.2in]
  \caption{$\chi$ plot for skew-$t$, and Gaussian}
  \end{figure}
\end{frame}

\begin{frame}{Extension of the skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Skew-$t$ distribution addresses two modeling concerns
    \begin{itemize}
      \item Extremal dependence
      \item Reasonable computing
    \end{itemize}
    \item Our contribution is to extend the skew-$t$
    \begin{itemize}
      \item Censoring to focus on extreme observations
      \item Partitioning to address long-range dependence
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Censoring data to focus on tail behavior}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item We censor the observed data at a high threshold $T$
    \item Censored data
    \begin{align*}
      \tilde{Y}_t(\bs) = \left\{ \begin{array}{ll}
          Y_t(\bs) \quad & \delta(\bs) = 1\\
          T \quad & \delta(\bs) = 0
      \end{array}\right.
    \end{align*}
    where $\delta(\bs) = I[Y(\bs) > T]$
    \item Allows tails of the distribution to speak for themselves
  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution properties}
  \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Model properties
    \begin{itemize}
    	\item Has flexible tail
      \begin{itemize}
        \item Skewness controlled by $\lambda$
        \item Weight of tails controlled by degrees of freedom $2a$
      \end{itemize}
    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011)
    	\item Computation that is on the order of Gaussian computation
    \end{itemize}
    \item Challenge: $\chi(h) > 0$ as $h \rightarrow \infty$ (Padoan, 2011)
    \begin{itemize}
      \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
      \item We deal with this through a daily random partition (similar to Kim et al., 2005 for non-extreme modeling)
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Example partition}
    \centering
    \begin{figure}
    \includegraphics[width=0.54\linewidth, trim=0 0 0 1in]{./plots/pot/example-partition-1.pdf}
    \includegraphics[width=0.54\linewidth, trim=0 0 0 1in]{./plots/pot/example-partition-2.pdf}
    \caption{Two sample partitions (number is at partition center)}
    \end{figure}
\end{frame}

\begin{frame}{$\chi(h)$ plot}
  \vspace{-2em}
  \centering
  \begin{figure}
  \includegraphics[width=.65\linewidth]{./plots/pot/chi-h.pdf}\\[-0.2in]
  \caption{$\chi$ plot for different data settings}
  \end{figure}
\end{frame}

\begin{frame}{Random partition skew-$t$ model}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item This new model is called a random partition skew-$t$ model, and it has the properties we desire
    \begin{itemize}
      \item Marginal distribution with flexible tails
      \begin{itemize}
        \item $\lambda$ term allows for asymmetry
        \item Degrees of freedom control heavy vs light tails
      \end{itemize}
      \item Asymptotic spatial dependence for that decays with distance between sites through partitioning
      \item Computation is on the order of Gaussian models for large space-time datasets
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{MCMC details}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Three main steps
    \begin{enumerate}[1.]
      \item Impute censored data below $T$
      \item Update parameters with standard random walk Metropolis Hastings or Gibbs sampling
      \item Make spatial predictions
    \end{enumerate}
    \item Priors are selected to be conjugate when possible
  \end{itemize}
\end{frame}

@@

\begin{frame}{Simulation study}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 50 dataset for each setting
    \begin{itemize}
      \item 144 sites in $[0, 10] \times [0, 10]$
      \begin{itemize}
        \item 100 training
        \item 44 testing
      \end{itemize}
    \end{itemize}
    \item Model parameters
    \begin{itemize}
      \item Spatial range: $\rho = 1$
      \item Skew parameter: $\lambda = 3$
      \item Degrees of freedom: 6 for $t$ distributions
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Simulation study results}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item Key findings
    \begin{itemize}
      \item Improvement over Gaussian methods when partitioning
      \item Specifying too few knots has a detrimental impact
      \item In all cases, non-thresholded models perform better than thresholded models
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Data analysis}
\begin{columns}[c]
\column{.45 \linewidth}
	\begin{itemize} \setlength{\itemsep}{0.5em}
	\item Ozone measurements
	\begin{itemize}
		\item max 8-hour ozone measurements
		\item data from 1089 sites
		\item July 2005
	\end{itemize}
  \item We take a stratified sample of $n = 800$ sites
  \begin{itemize}
    \item 271 from northeast
    \item 96 from northwest
    \item 269 from southeast
    \item 164 from southwest
  \end{itemize}
	\end{itemize}

@@

\begin{frame}{Data analysis}
  \centering
  \begin{figure}
    \includegraphics[width=\linewidth, trim=0 1in 0 0.7in ]{./plots/pot/ozone-10jul-us.pdf}
    \caption{Max 8-hour ozone measurements on July 10, 2005}
   \end{figure}
\end{frame}

\begin{frame}{Exploratory data analysis}
	\centering
  \begin{figure}
    \includegraphics[width=1\linewidth, trim=0 0 0 1in]{./plots/pot/chi-plot-ozone-res.pdf}
    \caption{$\widehat{\chi}$-plot for sample quantiles of ozone observations}
  \end{figure}
\end{frame}

\begin{frame}{Model comparisons}
  \begin{itemize} \setlength{\itemsep}{0.5em}
    \item 9 different analysis methods incorporating
    \begin{itemize}
      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
      \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
      \item 4 threshold levels
      \begin{itemize}
         \item $T = 0$
         \item $T = 50$ppb, $q(0.48)$
         \item $T = 75$ppb, $q(0.92)$
         \item $T = 85$ppb, $q(0.97)$
      \end{itemize}
    \end{itemize}
    \item Compare Brier scores using two-fold cross validation
  \end{itemize}
\end{frame}
